---
title: "Make hometown map art"
author: "Danielle Ferraro"
date: "`r Sys.Date()`"
output: pdf_document
---

# About

For the holidays this year, I wanted to give my parents a map of my hometown (heyo Bethel, CT) that could be printed and hung on a wall. This is what I came up with, using my beloved ggplot.

Some maps that inspired me:

- [](https://twitter.com/taraskaduk/status/1207295701983670274) 
- [](https://erdavis.com/2019/07/27/the-beautiful-hidden-logic-of-cities/)

# Setup

```{r}
library(here)
library(tidyverse)
library(sf)
library(tigris) # Data source
library(smoothr) 
library(showtext) # Fonts, depends on {sysfonts}
library(ggimage) # For geom_image()
library(colorspace) # For darken()
```

# Read in data

I used the [tigris](https://github.com/walkerke/tigris) package to directly download TIGER/Line shapefiles of town boundaries and other data for Fairfield county. Note: TIGER stands for Topologically Integrated Geographic Encoding and Referencing and is the format used by the US Census Bureau. There are a lot of different datasets from which you could download the same or similar data, such as OpenStreetMap. I poked around a few different data sources when starting this project, but found the `tigris` package had everything I needed and was simple to use. 
```{r}
# Census subdvisions, aka towns
fairfield_towns <- county_subdivisions(state = "CT", county = "Fairfield")

# Roads
fairfield_roads <- roads(state = "CT", county = "Fairfield")

# Rails
rails <- rails() # National dataset
fairfield_rails <- st_intersection(rails, fairfield_towns)

# Hydrology
fairfield_water_linear <- linear_water(state = "CT", county = "Fairfield")
fairfield_water_area <- area_water(state = "CT", county = "Fairfield")
```

I also defined a set of data points that I wanted to layer on top - the homes I lived in! I found a house icon to use to mark them on the map.
```{r}
our_places <- tibble(name = c("6 Terry Dr", "20 Cortland Dr"),
                     x = c(-73.394546, -73.382509),
                     y = c(41.387528, 41.403420)
)
our_places$img <- here("house_icon.png")
```

# Subset data

Subset Bethel town boundary, and define bounding box coordinates to crop the other data prior to plotting. The area will include Bethel with a small buffer around it. It is easier (faster) to crop the data now then to do so while plotting.
```{r}
bethel_boundary <- filter(fairfield_towns, NAME == "Bethel")

bbox <- st_bbox(st_buffer(bethel_boundary, 0.01))

bethel_roads <- st_crop(fairfield_roads, bbox)
bethel_water_linear <- st_crop(fairfield_water_linear, bbox)
bethel_water_area <- st_crop(fairfield_water_area, bbox)
bethel_rails <- st_crop(rails, bbox)
```

# Light data tweaking

Some of the water area polygons are kinda jagged. 
```{r}
plot(bethel_water_area[1])
```

I can smooth those polygons to make them a bit more aesthetically pleasing with the `smoothr::smooth()` function.
```{r}
bethel_water_area <- smooth(bethel_water_area, method = "chaikin")
plot(bethel_water_area[1])
```

# Plot

Choose fonts for the plot title and text labels. I learned you can download Google Fonts with the `sysfonts::font_add_google()` function, and manually browsed through them at [fonts.google.com](https://fonts.google.com/) til I found ones I liked.
```{r}
font_add_google(name = "Montserrat", family  = "Montserrat")
font_add_google(name = "Assistant", family = "Assistant")
showtext_auto() # Loads the fonts
```

Put it all together! In V2 I used `st_intersection()` and `st_difference()` right in the ggplot call to isolate data inside and outside of the town border, in order to color them differently.
```{r}
water_col <- "#7cc3c4" # bright turquoise
background_col <- "#2C3546" # deep gray/navy

# V1 - all roads/rails white
ggplot() +
  
  # Water
  geom_sf(data = bethel_water_area, fill = water_col, color = NA) +
  geom_sf(data = bethel_water_linear, color = water_col, size = 0.7) +

  # Roads - make main roads darker and thicker
  geom_sf(data = bethel_roads %>% filter(RTTYP %in% c("I", "S")), fill = NA, color = "white", size = 0.5) + 
  geom_sf(data = bethel_roads %>% filter(!RTTYP %in% c("I", "S")), fill = NA, color = "white", size = 0.2) + 
  
  # Rails
  geom_sf(data = bethel_rails, linetype = "twodash", size = 0.5, color = "white") +
    
  # Town border
  geom_sf(data = bethel_boundary, size = 1, fill = NA, color = "white") + 
  
  # Personal points
  geom_image(data = our_places, aes(x = x, y = y, image = img), size = 0.01, asp = 0.75, nudge_x = 0.00025) +
  geom_text(data = our_places, aes(x = x, y = y, label = name), size = 3, color = "white", family = "Assistant", fontface = "bold", hjust = -0.17) +
  #geom_label(data = our_places, aes(x = x, y = y, label = name), size = 3, label.size = NA, alpha = 0.1, color = "white", family = "Assistant", fontface = "bold", hjust = -0.15) +
  
  # Annotate sneaky signature in Huntington Pond
  annotate(geom = "text", x = -73.357900, y = 41.343654, label = "DMF\n2020", color = darken(water_col), size = 0.8, alpha = 0.4) +
  
  # Add labels and theme
  labs(caption = "BETHEL | CONNECTICUT") +
  theme_void() +
  theme(plot.caption = element_text(color = "white", size = 50, hjust = 0.5, family = "Montserrat"),
        plot.background = element_rect(fill = background_col),
        plot.margin = unit(c(0.25,0.77,0.5,0.77),"in")) # Temporary hack to make bg color fill the pdf


# V2 - with roads/rails outside Bethel grayed out 
ggplot() +
  
  # Water
  geom_sf(data = bethel_water_area, fill = water_col, color = NA) +
  geom_sf(data = bethel_water_linear, color = water_col, size = 0.7) +
  
  # Roads - make main roads darker and thicker
  geom_sf(data = st_intersection(bethel_roads, bethel_boundary) %>% filter(RTTYP %in% c("I", "S")), fill = NA, color = "white", size = 0.5) + 
  geom_sf(data = st_intersection(bethel_roads, bethel_boundary) %>% filter(!RTTYP %in% c("I", "S")), fill = NA, color = "white", size = 0.2) + 
  
  geom_sf(data = st_difference(bethel_roads, bethel_boundary) %>% filter(RTTYP %in% c("I", "S")), fill = NA, color = darken("white", 0.3), size = 0.5) + 
  geom_sf(data = st_difference(bethel_roads, bethel_boundary) %>% filter(!RTTYP %in% c("I", "S")), fill = NA, color = darken("white", 0.3), size = 0.2) +
  
  # Rails
  geom_sf(data = st_intersection(bethel_rails, bethel_boundary), linetype = "twodash", size = 0.5, color = "white") +
  geom_sf(data = st_difference(bethel_rails, bethel_boundary), linetype = "twodash", size = 0.5, color = darken("white", 0.3)) +
  
  # Town border
  geom_sf(data = bethel_boundary, size = 1, fill = NA, color = "white") + 
  
  # Personal points
  geom_image(data = our_places, aes(x = x, y = y, image = img), size = 0.01, asp = 0.75, nudge_x = 0.00025) +
  geom_text(data = our_places, aes(x = x, y = y, label = name), size = 3, color = "white", family = "Assistant", fontface = "bold", hjust = -0.17) +
  #geom_label(data = our_places, aes(x = x, y = y, label = name), size = 3, label.size = NA, alpha = 0.1, color = "white", family = "Assistant", fontface = "bold", hjust = -0.15) +
  
  # Annotate sneaky signature in Huntington Pond
  annotate(geom = "text", x = -73.357900, y = 41.343654, label = "DMF\n2020", color = darken(water_col), size = 0.8, alpha = 0.4) +
  
  # Add labels and theme
  labs(caption = "BETHEL | CONNECTICUT") +
  theme_void() +
  theme(plot.caption = element_text(color = "white", size = 50, hjust = 0.5, family = "Montserrat"),
        plot.background = element_rect(fill = background_col),
        plot.margin = unit(c(0.25,0.77,0.5,0.77),"in")) # Temporary hack to make bg color fill the pdf
```


# Save

```{r eval = FALSE}
ggsave(here("bethel_v2.pdf"), width = 12, height = 16, units = "in", dpi = "retina")
```